---
title: "Introduction to _CloseEnough_"
author: "Maxence Le Vasseur"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use this package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6
)
```

This package uses two-dimensional particle trajectories to study the distribution of particles relative to surrounding cellular structures. It calculates the nearest distance between each point of a trajectory and surrounding structures reprensented as thresholded binary images taken from a corresponding sequence of reference images. It can also be used to randomize the trajectories within a pre-defined area in order to compare the positioning of particles relative to other cellular structures to a stochastic distribution.

## Getting started

This R package that can be downloaded from [GitHub](https://github.com/MaxLev/CloseEnough). To install this package from GitHub, the _devtools_ package must first be installed. To install _devtools_, start R and enter:

```{r installation, eval=FALSE}
install.packages("devtools")
library(devtools)
```

Then to install the _CloseEnough_ enter:

```
install_github("MaxLev/CloseEnough")
```

Once the _CloseEnough_ package is installed, it can be loaded by the following command:

``` {r library, message = FALSE}
library(CloseEnough)
```

## Display images

This package depends on the _EBImage_ package to upload and display images. To start off, let us display an image distributed with the _CloseEnough_ package. 

```{r readImage}
## Load image
data(cellmovie)

## Display image
EBImage::display(cellmovie, method = "browser")
```

## Display cell outline and trajectories

Let us display the cell outline and overlay the particle trajectories on the image.

```{r display, message = FALSE}
## Load cell outline
data(outline)

## Load trajectory
data(trajectory)

## Display image
EBImage::display(cellmovie, method = "raster")

## Trace outline
lines(outline$x, outline$y, type = "l", col = "white")

## Draw trajectories
ntraj = nlevels(trajectory["trajectory"])
for (i in 1:ntraj) {
  traj_i <- subset(trajectory["data"], trajectory == i)
  lines(traj_i$x, traj_i$y, type="l", col = "white")
}

```

## Particle trajectories can be rotated or mirrored if needed
Let us rotate the particle trajectories to the right by a 90 degree angle:

``` {r, message = FALSE}
rotated_traj <- rotate_trajectory(trajectory, 90)
```

or reflect them along the y axis

``` {r, message = FALSE}
mirror_traj <- mirror_trajectory(trajectory, axis = "y")
```

Now let us visualize those different trajectories on a simple cell mask

``` {r, message = FALSE}
## Load cell mask
data(cellmask)

## Display image
EBImage::display(cellmask, method = "raster")

## Draw trajectories in black
ntraj = nlevels(trajectory["trajectory"])
for (i in 1:ntraj) {
  traj_i <- subset(trajectory["data"], trajectory == i)
  lines(traj_i$x, traj_i$y, type="l", col = "black")
}

## Add the rotated trajectories in blue
ntraj = nlevels(rotated_traj["trajectory"])
for (i in 1:ntraj) {
  traj_i <- subset(rotated_traj["data"], trajectory == i)
  lines(traj_i$x, traj_i$y, type="l", col = "blue")
}

## And the reflected trajectories in green
ntraj = nlevels(mirror_traj["trajectory"])
for (i in 1:ntraj) {
  traj_i <- subset(mirror_traj["data"], trajectory == i)
  lines(traj_i$x, traj_i$y, type="l", col = "green")
}

```

## Randomization of particle trajectories within the cytoplasm
_CloseEnough_ can easily randomize particle trajectories within a given area provided by a binary mask image. This provides valuable information regarding the spatial distribution of particles within a cellular context. As an example, let us randomize Panx2 foci trajectories within the cytoplasm and observe how their distribution changes relative to the mitochondrial network.


``` {r, message = FALSE}
## Randomize trajectories using an image mask as reference.
rand_traj <- randomize(trajectory, cellmask)

## Display image
EBImage::display(cellmovie, method = "raster")

## Trace cell outline
lines(outline$x, outline$y, type = "l", col = "white")

## Draw trajectories in white
ntraj = nlevels(trajectory["trajectory"])
for (i in 1:ntraj) {
  traj_i <- subset(trajectory["data"], trajectory == i)
  lines(traj_i$x, traj_i$y, type="l", col = "white")
}

## Add the randomized trajectories in blue
ntraj = nlevels(rand_traj["trajectory"])
for (i in 1:ntraj) {
  traj_i <- subset(rand_traj["data"], trajectory == i)
  lines(traj_i$x, traj_i$y, type="l", col = "blue")
}

```


## Calculate distribution probability
_CloseEnough_ can calculate the minimal distance between the particles and other cellular structures within a thresholded binary image. In this example, we will calculate the minimal distance between every single Panx2 foci used to build the trajectories and the mitochondrial network. We will also performed the same calculations with our randomized trajectories.

``` {r, fig.width = 8, fig.height = 6}
## Load the thresholded mask of the mitochondrial network
data(mitomask)

## Calculate minimal distance between Panx2 foci and mitochondria
# Channel refers to the channel that needs to be used within the binary image. In this case there is only one channel.
Panx2_mito <- min_distance(mitomask, trajectory, channel = 1)

## Calculate minimal distance between randomized Panx2 trajectories and mitochondria
rand_mito <- min_distance(mitomask, rand_traj, channel = 1)
```

Finally, these values can be visualized as probability distributions using any R graphic tools. The real and randomized probability distributions can also be directly compared with the function _profileplot_:

``` {r, message = FALSE}
graph <- profileplot(traj = Panx2_mito, randtraj = rand_mito)
## Display graph
graph
```
